---
title: "Task4"
author: "I Gusti Made Arisudana"
date: "4/27/2022"
output: html_document
---

```{r}
download.file("http://www.openintro.org/stat/data/ames.RData", destfile = "ames.RData")
load("ames.RData")
population <- ames$Gr.Liv.Area
samp <- sample(population, 60)
samp_mean <- mean(samp)
samp_sd <- sd(samp)
```

1.  Using the following function (which was downloaded with the data set), 
    plot all intervals. What proportion of your confidence intervals include 
    the true population mean? Is this proportion exactly equal to the 
    confidence level? If not, explain why.
```{r}
for(i in 1:50){
  samp <- sample(population, 60) 
  samp_mean[i] <- mean(samp)    
  samp_sd[i] <- sd(samp)    
}
lower_vector <- samp_mean - 1.96 * samp_sd / sqrt(60) 
upper_vector <- samp_mean + 1.96 * samp_sd / sqrt(60)
plot_ci(lower_vector, upper_vector, mean(population))
```
Berdasarkan hasil plotting diatas, terdapat 3 dari 50 interval yang tidak termasuk dalam nilai mu sebesar 1499,6904. Proporsinya adalah 94% dan nilainya tidak sama dengan nilai confidence level. Nilai confidence level dapat dianggap sebagai sebuah probabilitas mencakup true population mean.   
    
2.  Pick a confidence level of your choosing, provided it is not 95%. What is 
    the appropriate critical value?
```{r}
cl <- 97
alpha <- 1-(cl/100)
cp <- 1-(alpha/2) 
critical_value <- qnorm(cp)
critical_value
```
    
    
3.  Calculate 50 confidence intervals at the confidence level you chose in the 
    previous question. You do not need to obtain new samples, simply calculate 
    new intervals based on the sample means and standard deviations you have 
    already collected. Using the `plot_ci` function, plot all intervals and 
    calculate the proportion of intervals that include the true population 
    mean. How does this percentage compare to the confidence level selected for
    the intervals?
```{r}
for(i in 1:50){
  samp <- sample(population, 60) 
  samp_mean[i] <- mean(samp)    
  samp_sd[i] <- sd(samp)    
}
lower_vector <- samp_mean - critical_value * samp_sd / sqrt(60) 
upper_vector <- samp_mean + critical_value * samp_sd / sqrt(60)
plot_ci(lower_vector, upper_vector, mean(population))
```
Berdasarkan hasil plotting diatas 2 dari 50 interval tidak mencakup nilai mu = 1499,6904. Proporsinya adalah 96% dan nilainya lebih tinggi dari confidence level yang baru yaitu 97%.  