---
title: "Task7"
author: "I Gusti Made Arisudana"
date: "5/18/2022"
output: html_document
---

```{r}
download.file("http://www.openintro.org/stat/data/nc.RData", destfile = "nc.RData")
load("nc.RData")
```

## On your own

-   Calculate a 95% confidence interval for the average length of pregnancies 
(`weeks`) and interpret it in context. Note that since you're doing inference 
on a single population parameter, there is no explanatory variable, so you can 
omit the `x` variable from the function.
```{r}
inference(y = nc$weeks, est = "mean", type = "ci", null = 0,
alternative = "twosided", method = "theoretical", conflevel = 0.95)
```
Berdasarkan grafik diatas pada interval confidence 95%, didapatkan bahwa rata-rata dari lama waktu kehamilan berkisar dari 38,1528 - 38,5165 minggu.

-   Calculate a new confidence interval for the same parameter at the 90% 
confidence level. You can change the confidence level by adding a new argument 
to the function: `conflevel = 0.90`.
```{r}
inference(y = nc$weeks, est = "mean", type = "ci", method = "theoretical", conflevel = 0.90)
```
Berdasarkan grafik diatas pada interval confidence 90%, didapatkan bahwa rata-rata dari lama waktu kehamilan berkisar dari lama waktu hamil berkisar dari 38,182 - 38,4873 minggu.

-   Conduct a hypothesis test evaluating whether the average weight gained by 
younger mothers is different than the average weight gained by mature mothers.
```{r}
inference(y = nc$gained, x = nc$mature, est = "mean", type = "ht", null = 0, alternative = "twosided", method = "theoretical")
```
Berdasarkan perhitungan di atas, p-value memiliki nilai sebesar 0,1686 yang dimana nilai tersebut lebih besar dari nilai Alpha yang hanya sebesar 0,05. Dapat disimpulkan bahwa tidak ada perbedaan pertambahan berat badan rata-rata yang signifikan antara younger mom dan mature mom.

-   Now, a non-inference task: Determine the age cutoff for younger and mature 
mothers. Use a method of your choice, and explain how your method works.
```{r}
c(min(subset(nc, mature == 'younger mom')$mage),max(subset(nc, mature == 'younger mom')$mage))
c(min(subset(nc, mature != 'younger mom')$mage),max(subset(nc, mature == 'mature mom')$mage))
```
Dengan cara melakukan subset pada younger mom dan mature mom, didapat bahwa bahwa batas umur younger mom berkisar antara 13 hingga 34 tahun. Sedangkan untuk mature mom berkisar antara 35-50 tahun.


-   Pick a pair of numerical and categorical variables and come up with a 
research question evaluating the relationship between these variables. 
Formulate the question in a way that it can be answered using a hypothesis test
and/or a confidence interval. Answer your question using the `inference` 
function, report the statistical results, and also provide an explanation in 
plain language.
Research Question : Apakah ada perbedaan lama kehamilan berdasarkan usia?
```{r}
inference(nc$weeks, nc$mature, type="ht", est="mean", 
          null=0, method="theoretical", alternative="twosided")
```
Dari perhitungan diatas, didapati p-value sebesar 0.2258. Dapat disimpulkn bahwa tidak ada perbedaan yang signifikan antara rata-rata durasi kehamilan antara younger mom dan mature mom.
