---
title: "Task6"
author: "I Gusti Made Arisudana"
date: "5/18/2022"
output: html_document
---

```{r load-data, eval=FALSE}
download.file("http://www.openintro.org/stat/data/atheism.RData", destfile = "atheism.RData")
load("atheism.RData")
```

The question of atheism was asked by WIN-Gallup International in a similar 
survey that was conducted in 2005. (We assume here that sample sizes have 
remained the same.) Table 4 on page 13 of the report summarizes survey results 
from 2005 and 2012 for 39 countries.

-   Answer the following two questions using the `inference` function. As 
    always, write out the hypotheses for any tests you conduct and outline the
    status of the conditions for inference.

    **a.** Is there convincing evidence that Spain has seen a change in its 
    atheism index between 2005 and 2012?\
    *Hint:* Create a new data set for respondents from Spain. Form
    confidence intervals for the true proportion of athiests in both
    years, and determine whether they overlap.
```{r}
spain05 <- subset(atheism, nationality == "Spain" & year == "2005")
spain05$nationality <- as.factor(as.character(spain05$nationality))
table(spain05$nationality, spain05$response)

spain12 <- subset(atheism, nationality == "Spain" & year == "2012")
spain12$nationality <- as.factor(as.character(spain12$nationality))
table(spain12$nationality, spain12$response)
```
```{r}
inference(spain05$response, est = "proportion", type = "ci", method = "theoretical", success = "atheist")
inference(spain12$response, est = "proportion", type = "ci", method = "theoretical", success = "atheist")
```
H0 : Jumlah Atheists tidak berubah antara tahun 2005 dan 2012 di Spanyol
HA : Jumlah Atheists berubah antara tahun 2005 dan 2012 di Spanyol

Terdapat overlap yang signifikan antara interval cofident untuk sampel 2005 dan sampel 2012,  maka H0 (null hypothesis) tidak bisa diabaikan dan dapat disimpulkan bahwa tidak terdapat perubahan pada atheism index di Spanyol antara tahun 2005 dan 2012. Perubahan yang terjadi terhadap Atheisme berkemungkinan besar akibat kebetulan.

    **b.** Is there convincing evidence that the United States has seen a
    change in its atheism index between 2005 and 2012?
```{r}
us05 <- subset(atheism, nationality == "United States" & year == "2005")
us05$nationality <- as.factor(as.character(us05$nationality))
table(us05$nationality, us05$response)

us12 <- subset(atheism, nationality == "United States" & year == "2012")
us12$nationality <- as.factor(as.character(us12$nationality))
table(us12$nationality, us12$response)
```
```{r}
inference(us05$response, est = "proportion", type = "ci", method = "theoretical", success = "atheist")
inference(us12$response, est = "proportion", type = "ci", method = "theoretical", success = "atheist")
```
H0 : Jumlah Atheists tidak berubah antara tahun 2005 dan 2012 di Amerika
HA : Jumlah Atheists berubah antara tahun 2005 dan 2012 di Amerika

Tidak ada overlap antara interval confidence untuk sampel 2005 dan sampel 2012, maka H0 (Null Hypothesis) dapat diabaikan Perubahan Terhadap Atheism tidak mungkin terjadi karena kebetulan